Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-06-03 doc/paper.tex
scanned through my notes for anything else I might want to add

% TODO: follow the 300-word website limit as a guideline to abstract length? check pdf guide
\begin{abstract}
The large-scale generation of synthetic parallel corpora for data-driven machine translation is proposed.
A rudimentary system for doing this is presented, with mechanisms for syntactic complexity in the spirit of classical rule-based machine translation and high-quality sentence pairs in the spirit of example-based machine translation.
Preliminary experiments integrating these synthetic corpora into a shared task are reported.
Promising future directions are discussed, in particular, large-scale contributions from ``amateur linguists'' via crowdsourcing.
\end{abstract}

@@

As early as Jelinek's (in)famous quote during the 1980's that ``{\em Every time I fire a linguist, my performance goes up},'' both linguists and linguistics have been increasingly marginalized in natural language processing (NLP). 
More recently, for example, \mynewcite{collobert:2011} mention the possibility ``{\em [ideally]...to learn from letter sequences rather than words},'' illustrating an AI-motivated desire to move away from even the most fundamental of linguistic concepts.
A specific instance of this overall trend is the gradual shift of machine translation (MT) away from classical linguistic rule-based approaches \mycitep{hutchins:2005} to contemporary machine learning approaches, which underlie statistical phrase-based systems like Moses \mycitep{moses} and more recent neural network systems \mycitep{sutskever:2014,bengio:2014}.  
But in the end, even the most cutting-edge machine learning MT system relies crucially on multilingual parallel corpora, typically taken from external sources like multilingual parliamentary proceedings. 
This results in what we perceive to be a disproportionate amount of effort being dedicated to the machine learning component of machine translation, with input corpora being treated almost as an afterthought.\footnote{
    Of course, these parallel corpora represent hundreds of translator-hours of work. 
    However, these translations are generally optimized for human consumption, not machine learning, with no regard for such niceties as sentence alignment.
    \amtaonly{This is illustrated briefly in Section \ref{subsec:case_study}.}
    }

@@

One of the major advantages of data-driven MT is that it removes the need for linguists to develop a new set of rules for an MT system to support a new language.
In a sense, though, the rule-based MT linguist's job has really been  ``outsourced'' to third-party translators, who implicitly encode linguistic information in parallel corpora.
We believe there is still a significant ``in-house'' role for linguists to play in corpus-based machine translation, and that is to direct the crafting of high-quality, ``designer'' synthetic parallel corpora, leveraging technologies like crowdsourcing to maximize throughput.
Such a division of labor would let linguists be linguists, with MT system development proceeding in parallel.

@@

% TODO: don't need to go into this much detail, do you? This would be the first section to trim, if it runs OVER length
% To find an example of suboptimally translated parallel corpora in the literature, one need not look far.

As an example of suboptimal parallel corpora being used for machine translation, consider the official Chinese to English baseline translation of the {\small \tt tst2010} data set\footnote{https://wit3.fbk.eu/score.php?release=2014-01} from the TED talks corpus, described in greater detail in Section \ref{sec:experiments}.
As a very crude way of assessing translation quality, we used the official evaluation software MultEval \mycitep{multeval} to compute scores for each translated line individually.
Upon manual inspection, the worst scoring individual lines (with TER scores over 150\%) were almost all found to correspond to poor human translations on the Chinese side of the gold corpus.\footnote{
    High TER scores indicated misaligned sentences as well as clerical errors.  In particular, the lecture with {\small \tt talkid} 767 had dozens of erroneous lines, in some of the worst cases having the same Chinese sentence copy/pasted three times on the same line. 
    The baseline Moses system tries valiantly, but it ultimately scores poorly for trying to produce a good machine translation of a poor human translation.
    Alternatively, one can interpret these errors as an additional source of noise in the human encoding of English into Chinese.
}
Such translations might suffice for viewing translated videos monolingually, but they are far from ideal for training a machine translation system.

@@

% TODO: table with change; also list best and worst 
% caption should mention that the difference between 
    % ALTHOUGH you would expect that it would improve all models, it's not clear whether...err? hard to be precise here.
%Removing TED talk number 767 (over 200 lines), as well as 15 hand-picked lines with bad TER scores, 
%First, note that the scores improve, which is consistent with the observation that 
% something to keep in mind in case reviewers ask
%This results in a significant TER score change that is almost as large as the score difference between the highest and lowest ranking system from the evaluation campaign.  
%BLEU scores \mycitep{bleu}, on the other hand, did not change much.
% no, omit this, because DUH, if you change the test set, the scores will change. (it's like changing your ruler.)
% TODO: table and reference
%\begin{table}[]
%\centering
%\label{table:tst2010}
%\begin{tabular}{|l|l|l|}
%\hline
% & BLEU & TER  \\ \hline
%Full {\small \tt tst2010} & 11.2 & 77.0  \\ \hline
%Purged {\small \tt tst2010} & 11.8 & 71.3  \\ \hline
%%\hline
%%{\small \tt tst2014} &  &   \\ \hline
%% &  &   \\ \hline
%\end{tabular}
%\caption{By comparison, the difference between the baseline TER score and (not shown) } % TODO: caption
%\end{table}
% TODO: table of scores after purging. This is pretty crucial? why?
	% i think i THOUGHT this was crucial because i had to back up my claims concretely...
	% but the claim from the data is tenuous at best - of course the scores change if the test set changes! so WHAT?
	% BESIDES, I tried training and tuning with purged corpora, and the difference was minimal. it's quite possible that people just don't CARE
} % \amtaonly

@@

Other practical considerations shine a favorable light on parallel corpus generation.
As the inverse problem of translation, multilingual generation is also an easier task in some senses. 
For example, lexical ambiguity can be largely alleviated,\footnote{
    Ideally the generator would have {\em a priori} knowledge about whether it is writing a sentence pair about, say,  a (financial) bank or a (river) bank.
} and segmentation of languages like Chinese \mycitep{segmenter:2005,segmenter:2008} becomes a non-issue.
Parallel corpora also provide a natural avenue through which to interface contemporary MT systems with the countless linguist-years of expertise that have been invested into rule-based MT systems, as discussed in Section \ref{sec:related}.
And unlike the various competing MT methods, which are by nature competitive, corpora are inherently collaborative.

@@

% these links are too long for EMNLP. TODO: add them as footnotes for AMTA?
% http://www.fb10.uni-bremen.de/anglistik/langpro/kpml/genbank/generation-bank.html
% http://www.fb10.uni-bremen.de/anglistik/langpro/kpml/ genbank/R3b12-English/Docu/ENGLISH-nigel-exerciseset-mismatches-19981209/EX-SET-1.html
In terms of freely available multilingual natural language generators, the most well-developed one we could find was KPML \mycitep{kpml}.
But sentence writing is fairly involved because the focus seems to be on having a well-developed linguistic theory to explain the surface forms.
In contrast, our primary goal is to generate parallel corpora of surface forms, affording us a significant amount of flexibility as to the method of generation.

@@

@@

% TODO: looks like this is GOING to run over length. save some if not all of this for AMTA, which is going to run way UNDER length
%\subsection{Simplistic Linguistics}
\amtaonly{
We adopt an exceedingly simplistic view of language as nouns, verbs, and their modifications and elaborations.
Inspired by early work on transformations by \mynewcite{chomsky}, we focus on declarative indicative present active sentences in the hopes that they can be transformed into other forms in the long run. 
Similarly, although we enforce some semantic constraints, our system generally produces nonsensical sentences like Chomsky's celebrated ``{\em Colorless green ideas sleep furiously.}''
The fact that verb phrases can also serve as nouns (gerundives) and modifiers (participles) appears to be a common source of complexity, and we made it a priority to implement participial modifiers.
To capture recursive structures like participles, we adopt a rudimentary tree representation that is something intermediate between constituency trees and dependency trees, with constituent-like modifiers that also maintain dependency-like links to their targets.
}

@@

% TODO: should I bother with this?

@@

% running over length, but let's just write it all down and then see what to cut (hint: that stupid tst2010 stuff. maybe just summarize it in a footnote)

>>>>

% website limits abstract to 300-word, but that's WAY too long. Mine is about 70 words, and it's short but not too short.
\begin{abstract}
The large-scale generation of synthetic parallel corpora for data-driven machine translation is proposed.
A rudimentary system for doing this is presented, with mechanisms for syntactic complexity in the spirit of classical rule-based machine translation and high-quality sentence pairs in the spirit of example-based machine translation.
Preliminary experiments integrating these synthetic corpora into a shared task are reported.
Promising future directions are discussed, in particular, large-scale contributions from ``amateur linguists'' via crowdsourcing.
\end{abstract}

@@

As early as Jelinek's (in)famous quote during the 1980's that ``{\em Every time I fire a linguist, my performance goes up},'' both linguists and linguistics have been increasingly marginalized in natural language processing (NLP). 
More recently, for example, \mynewcite{collobert:2011} mention the possibility ``{\em [ideally]...to learn from letter sequences rather than words},'' illustrating an AI-motivated desire to move away from even the most fundamental of linguistic concepts.
A specific instance of this overall trend is the gradual shift of machine translation (MT) away from classical linguistic rule-based approaches \mycitep{hutchins:2005} to contemporary machine learning approaches, which underlie statistical phrase-based systems like Moses \mycitep{moses} and more recent neural network systems \mycitep{sutskever:2014,bengio:2014}.  
But in the end, even the most cutting-edge machine learning MT system relies crucially on multilingual parallel corpora, typically taken from external sources like multilingual parliamentary proceedings. 
This results in what we perceive to be a disproportionate amount of effort being dedicated to the machine learning component of machine translation, with input corpora being treated almost as an afterthought.\footnote{
    Of course, these parallel corpora represent hundreds of translator-hours of work. 
    However, these translations are generally optimized for human consumption, not machine learning, with no regard for such niceties as sentence alignment.
    \amtaonly{This is illustrated briefly in Section \ref{subsec:case_study}.}
    }
% TODO: modernization of corpus acquisition / bring machine learning to bear here too

@@

% TODO (AMTA padding?): it's rather like having an athlete with an Olympic-level modern training regimen, when it comes to diet/nutrition, they'll just eat whatever they can find on the streets.
% Long and rich tradition of decades, if not centuries of linguists, "chefs" - want to exploit this somehow

One of the major advantages of data-driven MT is that it removes the need for linguists to develop a new set of rules for an MT system to support a new language.
In a sense, though, the rule-based MT linguist's job has really been  ``outsourced'' to third-party translators, who implicitly encode linguistic information in parallel corpora.
We believe there is still a significant ``in-house'' role for linguists to play in corpus-based machine translation, and that is to direct the crafting of high-quality, ``designer'' synthetic parallel corpora, leveraging technologies like crowdsourcing to maximize throughput.
Such a division of labor would let linguists be linguists, with MT system development proceeding in parallel.  
% TODO: let machines be machines, do what machines do best

@@

% To find an example of suboptimally translated parallel corpora in the literature, one need not look far.
% TODO: summarize this in a footnote for emnlp? is there room?
As an example of suboptimal parallel corpora being used for machine translation, consider the official Chinese to English baseline translation of the {\small \tt tst2010} data set\footnote{https://wit3.fbk.eu/score.php?release=2014-01} from the TED talks corpus, described in greater detail in Section \ref{sec:experiments}.
As a very crude way of assessing translation quality, we used the official evaluation software MultEval \mycitep{multeval} to compute scores for each translated line individually.
Upon manual inspection, the worst scoring individual lines (with TER scores over 150\%) were almost all found to correspond to poor human translations on the Chinese side of the gold corpus.\footnote{
    High TER scores indicated misaligned sentences as well as clerical errors.  In particular, the lecture with {\small \tt talkid} 767 had dozens of erroneous lines, in some of the worst cases having the same Chinese sentence copy/pasted three times on the same line. 
    The baseline Moses system tries valiantly, but it ultimately scores poorly for trying to produce a good machine translation of a poor human translation.
    Alternatively, one can interpret these errors as an additional source of noise in the human encoding of English into Chinese.
}
Such translations might suffice for viewing translated videos monolingually, but they are far from ideal for training a machine translation system.

@@

% caption should mention that the difference between 
    % ALTHOUGH you would expect that it would improve all models, it's not clear whether...err? hard to be precise here.
%Removing TED talk number 767 (over 200 lines), as well as 15 hand-picked lines with bad TER scores, 
%First, note that the scores improve, which is consistent with the observation that 
% something to keep in mind in case reviewers ask
%This results in a significant TER score change that is almost as large as the score difference between the highest and lowest ranking system from the evaluation campaign.  
%BLEU scores \mycitep{bleu}, on the other hand, did not change much.
% no, omit this, because DUH, if you change the test set, the scores will change. (it's like changing your ruler.)
% table and reference
%\begin{table}[]
%\centering
%\label{table:tst2010}
%\begin{tabular}{|l|l|l|}
%\hline
% & BLEU & TER  \\ \hline
%Full {\small \tt tst2010} & 11.2 & 77.0  \\ \hline
%Purged {\small \tt tst2010} & 11.8 & 71.3  \\ \hline
%%\hline
%%{\small \tt tst2014} &  &   \\ \hline
%% &  &   \\ \hline
%\end{tabular}
%\caption{By comparison, the difference between the baseline TER score and (not shown) } 
%\end{table}
% TODO: table of scores after purging. This is pretty crucial? why?
	% i think i THOUGHT this was crucial because i had to back up my claims concretely...
	% but the claim from the data is tenuous at best - of course the scores change if the test set changes! so WHAT?
	% BESIDES, I tried training and tuning with purged corpora, and the difference was minimal. it's quite possible that people just don't CARE
} % \amtaonly

@@

% TODO: a reference to ``blessing of dimensionality'' for the inverse problem? how it's relatively easy to generate an exponentially large number of sentences?
Other practical considerations shine a favorable light on parallel corpus generation.
As the inverse problem of translation, multilingual generation is also an easier task in some senses. 
For example, lexical ambiguity can be largely alleviated,\footnote{
    Ideally the generator would have {\em a priori} knowledge about whether it is writing a sentence pair about, say,  a (financial) bank or a (river) bank.
}, segmentation of languages like Chinese \mycitep{segmenter:2005,segmenter:2008} becomes a non-issue, and there is the potential for generating annotated corpora with zero annotation error.
Parallel corpora also provide a natural avenue through which to interface contemporary MT systems with the countless linguist-years of expertise that have been invested into rule-based MT systems, as discussed in Section \ref{sec:related}.
And unlike the various competing MT methods, which are by nature competitive, corpora are inherently collaborative.

@@

% these links are too long for EMNLP. TODO: add them as footnotes for AMTA?
% http://www.fb10.uni-bremen.de/anglistik/langpro/kpml/genbank/generation-bank.html
% http://www.fb10.uni-bremen.de/anglistik/langpro/kpml/ genbank/R3b12-English/Docu/ENGLISH-nigel-exerciseset-mismatches-19981209/EX-SET-1.html
In terms of freely available multilingual natural language generators, the most well-developed one we could find was KPML \mycitep{kpml}.
But making sentences using KPML is a lengthy process, involving a well-developed linguistic theory to explain the surface forms.
In contrast, our primary goal is to generate large parallel corpora of surface forms quickly. 
This also affords us a significant amount of flexibility as to the method of generation.
% TODO: for our purposes, the linguistic correctness of a generator (hokey hidden structures) is less important than whether it generates well-translated sentences.

@@

% TODO: didn't mention how you could easily create infinitely many sentences like ``the number of cats is N'', but they'll be soo correlated that they have zero translation value

@@

% looks like this is GOING to run over length. save some if not all of this for AMTA, which is going to run way UNDER length
%\subsection{Simplistic Linguistics}
\amtaonly{
We adopt an exceedingly simplistic view of language as nouns, verbs, and their modifications and elaborations.
Inspired by early work on transformations by \mynewcite{chomsky}, we focus on declarative indicative present active sentences in the hopes that they can be transformed into other forms in the long run. 
Similarly, although we enforce some semantic constraints, our system generally produces nonsensical sentences like Chomsky's celebrated ``{\em Colorless green ideas sleep furiously.}''
The fact that verb phrases can also serve as nouns (gerundives) and modifiers (participles) appears to be a common source of complexity, and we made it a priority to implement participial modifiers.
To capture recursive structures like participles, we adopt a rudimentary tree representation that is something intermediate between constituency trees and dependency trees, with constituent-like modifiers that also maintain dependency-like links to their targets.
}

@@

@@
